---
title: "Assignment: AI-assisted Code Reviews"
author: "Xingrong Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Assignment: AI-assisted Code Reviews

DAG 2:
dag {
bb="0,0,1,1"
AI [pos="0.200,0.700"]
Age [pos="0.200,0.200"]
Complexity [pos="0.600,0.350"]
Extent [pos="0.600,0.200"]
Quality [pos="0.600,0.500"]
Result [pos="0.800,0.700"]
Skill [pos="0.400,0.200"]
AI -> Extent
AI -> Quality
AI -> Result
Age -> AI
Age -> Skill
Complexity -> AI
Complexity -> Extent
Complexity -> Quality
Extent -> Result
Quality -> Result
Skill -> AI
Skill -> Extent
Skill -> Quality
}

```{r descriptives, message=FALSE, warning=FALSE}
library(tidyverse)

reviewers <- read.csv("Assignments/reviewers.csv", stringsAsFactors = FALSE)
reviews <- read.csv("Assignments/reviews.csv", stringsAsFactors = FALSE)
str(reviewers)
str(reviews)
head(reviewers)
head(reviews)
```

```{r merge-data, message=FALSE, warning=FALSE}
d <- reviews %>%
  left_join(reviewers, by = "reviewer.id")

glimpse(d)
```

## Task 2: Descriptive statistics and visualizations
```{r analyze-data, message=FALSE, warning=FALSE}
# 1) Distribution of AI usage (U)
ggplot(d, aes(x = factor(used.cr.technology))) +
  geom_bar() +
  labs(
    x = "AI usage (0 = no, 1 = yes)",
    y = "Number of reviews",
    title = "Distribution of AI-assisted vs non-AI reviews"
  )
```

```{r analyze-data-2, message=FALSE, warning=FALSE}
# 2) Distribution of extent (E)
ggplot(d, aes(x = extent)) +
  geom_histogram(bins = 30) +
  labs(
    x = "Extent (number of characters)",
    y = "Count",
    title = "Distribution of review extent"
  )
```

```{r analyze-data-3, message=FALSE, warning=FALSE}
# 3) Distribution of quality (Q)
ggplot(d, aes(x = factor(quality))) +
  geom_bar() +
  labs(
    x = "Quality rating (1–5)",
    y = "Number of reviews",
    title = "Distribution of expert quality ratings"
  )
```

```{r analyze-data-4, message=FALSE, warning=FALSE}
# 4) Optional: Relationship between complexity and extent, without regression line
ggplot(d, aes(x = complexity, y = extent)) +
  geom_point(alpha = 0.4) +
  labs(
    x = "Complexity (standardized)",
    y = "Extent (characters)",
    title = "Complexity vs. review extent (descriptive scatter)"
  )
```

## Task 3: Bayesian regression models
```{r}
library(rethinking)

# Standardize background variables
d$age_z   <- scale(d$age)
d$skill_z <- scale(d$skill)
d$comp_z  <- scale(d$complexity)
```

```{r models, message=FALSE, warning=FALSE}
# Exponential model for extent
dat_extent <- list(
  E = d$extent,
  U = d$used.cr.technology,
  A = as.numeric(d$age_z),
  S = as.numeric(d$skill_z),
  C = as.numeric(d$comp_z),
  N = nrow(d)
)

m_extent <- ulam(
  alist(
    E ~ dexp(lambda),
    log(lambda) <- aE + bEU*U + bEA*A + bES*S + bEC*C,
    aE  ~ dnorm(0, 1),
    c(bEU, bEA, bES, bEC) ~ dnorm(0, 0.5)
  ),
  data = dat_extent,
  chains = 4, cores = 4
)
```
All 4 chains finished successfully.
Mean chain execution time: 10.0 seconds.
Total execution time: 10.5 seconds.

```{r models-2, message=FALSE, warning=FALSE}
# Ordered logistic model for quality
dat_quality <- list(
  Q = d$quality,
  U = d$used.cr.technology,
  A = as.numeric(d$age_z),
  S = as.numeric(d$skill_z),
  C = as.numeric(d$comp_z),
  N = nrow(d)
)

m_quality <- ulam(
  alist(
    Q ~ dordlogit(phi, cutpoints),
    phi <- aQ + bQU*U + bQA*A + bQS*S + bQC*C,
    aQ  ~ dnorm(0, 1),
    c(bQU, bQA, bQS, bQC) ~ dnorm(0, 0.5),
    cutpoints ~ dnorm(0, 1.5)
  ),
  data = dat_quality,
  chains = 4, cores = 4
)
```
All 4 chains finished successfully.
Mean chain execution time: 787.7 seconds.
Total execution time: 804.6 seconds.

## Task 4: Prior predictive checks
```{r prior-check, message=FALSE, warning=FALSE}
# PRIOR predictive for extent
# m_extent_prior <- ulam(
#   alist(
#     E ~ dexp(lambda),
#     log(lambda) <- aE + bEU*U + bEA*A + bES*S + bEC*C,
#     aE  ~ dnorm(0, 1),
#     c(bEU, bEA, bES, bEC) ~ dnorm(0, 0.5)
#   ),
#   data = dat_extent,
#   chains = 1, cores = 1,
#   sample_prior = TRUE,
#   iter = 2000
# )

mu0 <- mean(d$extent)
m_extent_prior <- ulam(
  alist(
    E ~ dexp(lambda),
    lambda <- 1/mu,
    log(mu) <- a_mu + bU*U + bA*A + bS*S + bC*C,
    a_mu ~ dnorm(log(mu0), 0.5),
    c(bU, bA, bS, bC) ~ dnorm(0, 0.5)
  ),
  data = dat_extent,
  chains = 1, cores = 1,
  sample_prior = TRUE,
  iter = 2000
)

# Simulate prior predictive extent
yrep_extent <- sim(m_extent_prior, data = dat_extent, n = 1000)
yrep_extent <- as.vector(yrep_extent)
# apply(yrep_extent, 1, median) |> hist(main="Prior predictive medians")

par(mfrow = c(1,2))
hist(d$extent, breaks = 30, main = "Observed extent", xlab = "extent (characters)")
hist(yrep_extent, breaks = 30, main = "Prior predictive extent", xlab = "extent (characters)")
par(mfrow = c(1,1))
```

```{r prior-check-2, message=FALSE, warning=FALSE}
# PRIOR predictive for quality (ordered logistic)
m_quality_prior <- ulam(
  alist(
    Q ~ dordlogit(phi, cutpoints),
    phi <- aQ + bQU*U + bQA*A + bQS*S + bQC*C,
    aQ  ~ dnorm(0, 1),
    c(bQU, bQA, bQS, bQC) ~ dnorm(0, 0.5),
    cutpoints ~ dnorm(0, 1.5)
  ),
  data = dat_quality,
  chains = 1, cores = 1,
  sample_prior = TRUE,
  iter = 2000
)

# Prior predictive draws for quality (ordinal 1..5)
yrep_quality <- sim(m_quality_prior, data = dat_quality, n = 2000)
yrep_quality <- as.vector(yrep_quality)

# Compare marginal distributions (Observed vs Prior predictive)
obs_tab   <- table(factor(d$quality, levels = 1:5))
prior_tab <- table(factor(yrep_quality, levels = 1:5))

obs_prop   <- obs_tab / sum(obs_tab)
prior_prop <- prior_tab / sum(prior_tab)

barplot(
  rbind(obs_prop, prior_prop),
  beside = TRUE,
  legend.text = c("Observed", "Prior predictive"),
  names.arg = 1:5,
  main = "Quality: observed vs prior predictive",
  xlab = "quality (1–5)",
  ylab = "proportion"
)
```

## Task 5: Model diagnostics and posterior predictive checks
```{r diagnostics, message=FALSE, warning=FALSE}
# Summaries
precis(m_extent, 2)
precis(m_quality, 2)

# Trace plots
traceplot(m_extent)
traceplot(m_quality)

# stanfit extraction (works for most ulam models)
sf_extent  <- m_extent$stanfit
sf_quality <- m_quality$stanfit

# Rhat and neff ratio
max(rhat(sf_extent))
min(neff_ratio(sf_extent))

max(rhat(sf_quality))
min(neff_ratio(sf_quality))

```
max(rhat(m_extent$stanfit))
min(neff_ratio(m_extent$stanfit))

```{r}
library(rstan)

sum(sapply(get_sampler_params(m_extent$stanfit, inc_warmup = FALSE),
           function(x) x[,"divergent__"]))

sum(sapply(get_sampler_params(m_quality$stanfit, inc_warmup = FALSE),
           function(x) x[,"divergent__"]))

# # Check Rhat and n_eff manually
# library(coda)
# extent_chains  <- extract.samples(m_extent, as.matrix = TRUE)
# quality_chains <- extract.samples(m_quality, as.matrix = TRUE)

# # rethinking already prints Rhat and n_eff in precis(); use that in your report.
# max(rhat(m_extent@stanfit))
# min(neff_ratio(m_extent@stanfit))

# max(rhat(m_quality@stanfit))
# min(neff_ratio(m_quality@stanfit))
```

```{r}
# Numerical diagnostics
precis(m_extent, depth = 2)

# Trace plots
traceplot(m_extent)

# Numerical diagnostics
precis(m_quality, depth = 2)

# Trace plots
traceplot(m_quality)
```