# Exercises 1

install.packages(c("coda","mvtnorm","devtools","loo","dagitty","shape"))
devtools::install_github("stan-dev/cmdstanr")
cmdstanr::install_cmdstan()
devtools::install_github("rmcelreath/rethinking")

---

## 2.
Recall the globe tossing model from the chapter. 
Compute and plot the grid approximate posterior distribution 
for each of the following sets of observations. 
In each case, assume a uniform prior for p.


```{r}
# Create 100 evenly spaced values from 0 to 1
p_grid <- seq( from=0 , to=1 , length.out=100 ) 
```

1. W,W,W
```{r}
# likelihood of 3 water in 3 tosses
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

2. W,W,W,L
```{r}
# likelihood of 3 water in 4 tosses
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

3. L,W,W,L,W,W,W
```{r}
# likelihood of 5 water in 7 tosses
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

## 3.
Now assume a prior for p that is equal to zero when
p < 0.5 and is a positive constant when p >= 0.5. Again compute and
plot the grid approximate posterior distribution for each of the sets of
observations in the problem just above.

1. W,W,W
```{r}
p_grid <- seq( from=0 , to=1 , length.out=100 )
# likelihood of 3 water in 3 tosses
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

2. W,W,W,L
```{r}
# likelihood of 3 water in 4 tosses
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

3. L,W,W,L,W,W,W
```{r}
# likelihood of 5 water in 7 tosses
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```