---
title: "Exercise 1"
author: "Rona Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Exercises 1

install.packages(c("coda","mvtnorm","devtools","loo","dagitty","shape"))
devtools::install_github("stan-dev/cmdstanr")
cmdstanr::install_cmdstan()
devtools::install_github("rmcelreath/rethinking")

---

## 2.
Recall the globe tossing model from the chapter. 
Compute and plot the grid approximate posterior distribution 
for each of the following sets of observations. 
In each case, assume a uniform prior for p.

```{r}
# Create 100 evenly spaced values from 0 to 1
p_grid <- seq( from=0 , to=1 , length.out=100 ) 
```

1. W,W,W
```{r}
# likelihood of 3 water in 3 tosses
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

2. W,W,W,L
```{r}
# likelihood of 3 water in 4 tosses
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

3. L,W,W,L,W,W,W
```{r}
# likelihood of 5 water in 7 tosses
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
prior <- rep( 1 , 100 ) # uniform prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

## 3.
Now assume a prior for p that is equal to zero when
p < 0.5 and is a positive constant when p >= 0.5. Again compute and
plot the grid approximate posterior distribution for each of the sets of
observations in the problem just above.

1. W,W,W
```{r}
p_grid <- seq( from=0 , to=1 , length.out=100 )
# likelihood of 3 water in 3 tosses
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

2. W,W,W,L
```{r}
# likelihood of 3 water in 4 tosses
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

3. L,W,W,L,W,W,W
```{r}
# likelihood of 5 water in 7 tosses
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
prior <- ifelse(p_grid < 0.5, 0, 1) # modified prior
posterior <- likelihood * prior
posterior <- posterior / sum(posterior) # standardize
plot(posterior ~ p_grid, type="l")
```

## 4. Q&A

1. What does it mean when we say “All models are wrong, but some are useful”?

All models simplify reality, so they are never 100 % correct, 
but they can still help us reason, predict, and learn about the world.

2. What’s the difference between the frequentist and Bayesian view of probability?

Frequentist: probability = long-run frequency of events.
Bayesian: probability = degree of belief about something (how likely we think it is).

3. In Bayesian thinking, what does the prior represent?

The prior represents what we believe about a parameter (like p) before we see any data.

4. What is the posterior, and how is it related to the likelihood and prior?

Posterior ∝ Likelihood × Prior.
It is proportional to the likelihood (how well the data supports different parameter values) times the prior.
The posterior is our updated belief about the parameter after seeing the data.

5. what does each number mean, and what is the result?
dbinom(3, size = 3, prob = 0.5)

This gives the probability of getting exactly 3 successes (e.g., heads) in 3 trials (e.g., coin flips) 
when the probability of success on each trial is 0.5. The result is 0.125 (or 12.5%).

6. Why do we multiply likelihood * prior in Bayesian updating?

Because Bayes’ rule says that our new belief (posterior) is proportional to how well each hypothesis fits the data (likelihood) times how strongly we believed it before (prior).
The multiplication combines old knowledge with new evidence.

7. After we multiply likelihood and prior, why do we divide by sum(posterior)?

To make the posterior a proper probability distribution—it must sum (or integrate) to 1.
Dividing by sum(posterior) rescales the values so that all probabilities add up to 1.

8. If you observed 3 heads out of 3 flips, what kind of p values will have the highest posterior probability?

Large p values (close to 1).
Those values make “3 heads in 3 flips” very likely, so the posterior curve peaks near p = 1.

9. If you observed 1 head and 2 tails, how will the posterior shape change compared to 3 heads?

It would shift toward smaller p values (around 0.3–0.4) because those make “1 out of 3” more likely.
The distribution becomes lower near p = 1 and higher near p = 0.

10. Why is Bayesian inference sometimes described as “learning from data”?

Because we start with prior beliefs and update them using observed data to form a new belief (posterior).
Each dataset teaches the model something new.

11. What does it mean to “evaluate the model” in Bayesian workflow (design → condition → evaluate)?

After fitting the model, we check whether it makes realistic predictions—using posterior predictive checks or comparison with actual data—to see if our assumptions were reasonable.

12. Suppose your model’s predictions don’t match real data.
What might that tell you about your prior or likelihood assumptions?

Something is wrong with our assumptions—maybe the prior is unrealistic or the likelihood (data-generation story) is too simple.
We should redesign the model or choose different priors.


R hat, the trace box,