---
title: "Assignment: AI-assisted Code Reviews"
author: "Xingrong Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Assignment: AI-assisted Code Reviews

## Causal Assumptions
Outcomes:
- `extent (E)`: Number of characters in the code review.
- `quality (Q)`: Subjective quality of the code review as assessed by experts (where 1 is bad and 5 good quality).
- `result (R)`:Whether the code review contributed to making a decision on the pull request (1) or not (0).

Predictors:
- `age (G)`: Age of the reviewer in number of years.
- `skill (S)`: Skill of a reviewer measured on a continuous scale.
- `used.cr.technology (A)`: Whether the reviewer uses AI-support during code review (1) or not(0).
- `complexity (C)`: Relative complexity of the code change to be reviewed. 

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Load libraries
library(dagitty)
library(ggdag)
library(ggplot2)
library(rethinking)
library(dplyr)
library(tidyverse)
library(skimr)
theme_set(theme_dag_blank())
```

Make sure using the same **R session** that has:
```{r}
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

```{r data-prep}
reviewers <- read.csv("Assignments/reviewers.csv")
str(reviewers) # check format
reviewers <- as.list(reviewers) # make to a list so we can add alpha below

reviews <- read.csv("Assignments/reviews.csv")
str(reviews) # check format
reviews <- as.list(reviews) # make to a list so we can add alpha below

# Merge datasets
df <- merge(reviews, reviewers, by = "reviewer_id")
glimpse(df)

# Rename variables for Stan compatibility
df <- df %>%
    rename(
        used_cr_technology = used.cr.technology,
    )

# Standardize predictors
df <- df %>%
    mutate(
        age_s = scale(age),
        skill_s = scale(skill),
        complexity_s = scale(complexity),
        extent_s = scale(extent),
        quality_s = scale(quality)
    )

str(df)
```

### DAG 1：
Selection + Confounding + Mediation
- `age` -> `skill` (older reviewers are usually more experienced and skilled)
- `age` -> `used.cr.technology` (older reviewers are less likely to adopt AI)
- `skill` -> `extent` & `quality` (more skilled reviewers write more concise and higher quality reviews)
- `complexity` -> `extent` & `quality `(more complex code changes require longer reviews and harder to review well)
- `used.cr.technology` -> `extent` & `quality` (AI-assisted can change length and clarity of reviews)
- `extent` & `quality` -> `result`,  (more concise and higher quality reviews are more likely to have higher approval rates)
Aversion: `used.cr.technology` → `result` (direct negative path, “categorically discarded more often”)

Predictors (independent variables): `used.cr.technology`
Confounders: `age`, `skill`, `complexity`
Mediators: `extent`, `quality`
Outcomes (dependent variables): `result`

Therefore, 
- `age`, `skill`, and `complexity` are confounders that influence both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.
- `extent` and `quality` are mediators that lie on the causal path between the predictor `used.cr.technology` and the outcome `result`.

This DAG supports total vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.

age ─▶ skill ─▶ extent ─▶ result
 │             └▶ quality ─▶ result
 └────────▶ used.cr.technology ─▶ extent, quality ─▶ result
complexity ─▶ extent, quality, result

```{r dag1, fig.height=5, fig.width=7}
dag1 <- dagitty("dag {
    G -> S
    G -> A
    S -> E
    S -> Q
    C -> E
    C -> Q
    A -> E
    A -> Q
    E -> R
    Q -> R
    A -> R
}")

ggdag(dag1, text = FALSE, layout = "nicely") +
geom_dag_edges(edge_colour = "grey30") +
geom_dag_point(shape = 21, size = 16, fill = "#8ECAE6", colour = "#4E9AC3", stroke = 0.8) +
geom_dag_text(aes(label = name), size = 4, colour = "black") +
labs(title = "DAG 1: Selection + Confounding + Mediation") +
coord_equal()
```

### DAG 2：
Skill-driven; AI mainly a proxy for skill & complexity
adoption + Confounding + Mediation
- `age` -> `skill` (older reviewers are usually more experienced and skilled)
- `skill` -> `used.cr.technology` (more skilled reviewers are usually curious or confident enough to try new tools.)
- `skill` -> `extent` & `quality` (more skilled reviewers write more concise and higher quality reviews)
- `complexity` -> `extent` & `quality` (more complex code changes require longer reviews and harder to review well)
- `extent` & `quality` -> `result`,  (more concise and higher quality reviews are more likely to have higher approval rates)
`used.cr.technology` -> `result` (weak/zero) direct path to Result; main AI influence is mediated by Extent/Quality

Allow that `used.cr.technology` is a proxy for skill and complexity, and the main influence of AI on `result` is mediated by `extent` and `quality`.
most of the observed difference between AI and non-AI reviews is due to differences in skill and complexity, not a direct effect of AI.
Skill is a confounder that influences both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.

The apparent correlation between “AI users have lower approval rate” might not mean AI causes worse results.
It could mean “AI users are younger, less experienced, or handle harder reviews.”

After adjusting for those confounders (Skill, Complexity, Age), the direct effect of AI may disappear.

This DAG supports total vs adjusted vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
- Adjusted effect of `used.cr.technology` on `result` conditioning on confounders `age`, `skill`, and `complexity` but not mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.

age ─▶ skill ─▶ used.cr.technology ─▶ extent ─▶ result
 │                     │                └▶ quality ─▶ result
 │                     └────────────────────────────────▶ result (weak)
complexity ─────────────────────────▶ extent, quality, result

```{r dag2, fig.height=5, fig.width=7}
dag2 <- dagitty("dag {
    G -> S
    S -> A
    S -> E
    S -> Q
    C -> E
    C -> Q
    E -> R
    Q -> R
    A -> R
}")

ggdag(dag2, text = FALSE, layout = "nicely") +
geom_dag_edges(edge_colour = "grey30") +
geom_dag_point(shape = 21, size = 16, fill = "#BDEFB6", colour = "#69BE63", stroke = 0.8) +
geom_dag_text(aes(label = name), size = 4, colour = "black") +
labs(title = "DAG 2: Skill-driven; AI as proxy for skill & complexity") +
coord_equal()
```

```{r adjustment-sets}
adjustmentSets(dag1, exposure = "used_cr_technology", outcome = "result")
adjustmentSets(dag2, exposure = "used_cr_technology", outcome = "result")
```

Viability discussion. 
Compare which back-door paths need blocking and what each adjustment set implies (e.g., in DAG 1, include Age, Skill, Complexity to estimate AI’s effect; if you include Extent/Quality you move from total to direct effect). 
Tie your reasoning to the collider/mediator logic practiced in E3 (information, K–L, thinking carefully about model vs. data) and E5 (mediation pipes).

## Descriptive Statistics & Visualizations

Follow the standardize/visualize patterns from E2/E4:

Summaries and correlations (extent/quality/complexity/age/skill/AI/result).

Visuals: scatter (Extent vs Complexity), outcome by AI (box/scatter), approval rates by AI.

Normalize extent/quality where helpful; keep raw scales visible for interpretability.

```{r descriptives, fig.height=5, fig.width=7}
reviewers <- read_csv("Assignments/reviewers.csv")
reviews   <- read_csv("Assignments/reviews.csv")

# inspect
head(reviewers)
head(reviews)
```

```{r}
df <- reviews %>%
  left_join(reviewers, by = "reviewer.id")

# quick structure check
glimpse(df)

```

```{r}
df <- df %>%
    mutate(
        ai_user = as.integer(used.cr.technology),
        approved = as.integer(result == 1)
    )
```

```{r}
skim(df %>% select(extent, quality, complexity, age, skill))
```

## Statistical Modeling

### Option (a): 1 DAG → 1 model per outcome

Extent (continuous): linear model (or Student-t if heavy tails):

Extent_𝑖 ∼ 𝑁 (𝜇_𝑖, 𝜎), 𝜇_𝑖 = 𝛼 + 𝛽_AI * AI_𝑖 + 𝛽_𝑐 * Complexity_𝑖 + 𝛽_𝑎 * Age_𝑖 + 𝛽_𝑠 * Skill_𝑖.

(E2 shows linear regression workflow; E4 suggests adding varying intercepts by reviewer if you want a multilevel version.)


Quality (ordered 1–5): ordered logistic/probit:

Pr(Quality_𝑖 ≤ 𝑘) = logit*(−1) * (𝜏_𝑘 − 𝜂_𝑖), 𝜂_𝑖 = 𝛾_0 + 𝛾_AI * AI_𝑖 + 𝛾_𝑐 * Complexity_𝑖 + 𝛾_𝑎 * Age_𝑖 + 𝛾_𝑠 * Skill_𝑖.

(Ordered categorical modeling is exactly what E6 trains, including priors on cutpoints.)


Result (binary): logistic:

Result_𝑖 ∼ Bernoulli(𝑝_𝑖), logit(𝑝_𝑖) = 𝛿_0 + 𝛿_AI * AI_𝑖 + 𝛿_𝑐 * Complexity_𝑖 + 𝛿_𝑎 * Age_𝑖 + 𝛿_𝑠 * Skill_𝑖 + [𝛿_𝑒 * Extent_𝑖 + 𝛿_𝑞 * Quality_𝑖],

bracketed terms included when you target the direct (not total) effect of AI (E5).
Add random intercepts for reviewer.id to acknowledge repeated reviews (E4).

### Option (b): 1 outcome across ≥2 DAGs (recommended: Result)

Model R₁ (DAG 1, total effect): logit(Result) ~ AI + Age + Skill + Complexity.

Model R₂ (DAG 1, direct effect): logit(Result) ~ AI + Age + Skill + Complexity + Extent + Quality.
Compare R₁ vs R₂ (WAIC/LOO). If 𝛿_AI shrinks a lot in R₂, the effect is mediated by Extent/Quality (or by the aversion path you posit). (E5 gives the template for this comparison.)

## Priors & Prior Predictive Checks
Use standardized inputs to simplify priors (E2/E4/E6 style):

Intercepts (logit/ordered): Normal(0,1) or Normal(−1,1) when you believe baseline approval < 50% (see NWO example).

Slopes: Normal(0, 0.5) or Normal(0, 1) on standardized predictors (regularizing as in Wines2012).

Cutpoints (ordered quality): ordered priors (e.g., Normal with increasing constraints). (E6)

Group SDs (if multilevel): half-Normal(0, 1) on random-effect scales (E4).
Run prior predictive checks to ensure plausible ranges before training (E4/E6).

## Assess Training & Compare Models
Convergence (R-hat, trace plots), effective sample sizes (if Bayesian; E4).

LOO/WAIC to compare candidate models and decide which DAG mapping predicts best (E4/E7).

## Conclusions
Report causal statements with the adjustment set tied to your DAG.

Contrast total vs direct effect of AI on Result.

Discuss bias/aversion (direct AI → Result path) and whether differences remain after adjusting for Extent/Quality/Complexity.
This is exactly what the assignment asks you to synthesize at the end.