---
title: "Exercise 2"
author: "Rona Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Exercises 2

## 1.
The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. 
Provide predicted heights and 89% compatibility intervals for each of these individuals. 
That is, fill in the table below, using model-based predictions.

individual weight expected height 89% interval
1 46.95
2 43.72
3 64.78
4 32.59
5 54.63

Fit the model with quap, to provide a quadratic approximation of the posterior.

```{r}
library(rethinking)
data("Howell1")
d <- Howell1
d2 <- d[d$age >= 18, ]
xbar <- mean(d2$weight)
m <- quap(
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + b * (weight - xbar),
        a ~ dnorm(178, 20),
        b ~ dnorm(0, 10),
        sigma ~ dunif(0, 50)
    ),
    data = d2
)
```

Then produce samples from the quadratic approximate posterior.
You could use mvrnorm directly, or the convenient extract.samples
function:

```{r}
post <- extract.samples(m)
str(post)
```

```{r}
y <- rnorm( 1e5 , post$a + post$b*( 46.95 - xbar ) , post$sigma )
mean(y)
PI(y,prob=0.89)
```

```{r}
f <- function( weight ) {
y <- rnorm( 1e5 , post$a + post$b*( weight - xbar ) , post$sigma )
return( c( mean(y) , PI(y,prob=0.89) ) )
}
weight_list <- c(46.95,43.72,64.78,32.59,54.63)
result <- sapply( weight_list , f )
```

```{r}
rtab <- cbind( weight_list , t( result ) )
colnames(rtab) <- c("weight","height","5%","94%")
rtab
```


## 2.
Select out all the rows in the Howell1 data with ages
below 18 years of age. If you do it right, you should end up with a
new data frame with 192 rows in it.

(a) Fit a linear regression to these data, using quap. Present and
interpret the estimates. For every 10 units of increase in weight, how
much taller does the model predict a child gets?
```{r}
library(rethinking)
data("Howell1")
d <- Howell1
d3 <- d[d$age < 18, ] # only <18 yrs old
str(d3)
```

```{r}
xbar <- mean( d3$weight )
m <- quap(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b*( weight - xbar ) ,
a ~ dnorm( 178 , 20 ),
b ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 50 )
) , data=d3 )
precis(m)
```

(b) Plot the raw data, with height on the vertical axis and weight
on the horizontal axis. Superimpose the MAP regression line and
89% interval for the mean. Also superimpose the 89% interval for
predicted heights.
```{r}
post <- extract.samples( m )
w.seq <- seq(from=1,to=45,length.out=50)
mu <- sapply( w.seq , function(z) mean( post$a + post$b*(z-xbar) ) )
mu.ci <- sapply( w.seq , function(z)
PI( post$a + post$b*(z-xbar) , prob=0.89 ) )
pred.ci <- sapply( w.seq , function(z)
PI( rnorm( 10000 , post$a + post$b*(z-xbar) , post$sigma) , 0.89 ) )
```

```{r}
plot( height ~ weight , data=d3 ,
col=col.alpha("slateblue",0.5) , cex=0.5 )
lines( w.seq , mu )
lines( w.seq , mu.ci[1,] , lty=2 )
lines( w.seq , mu.ci[2,] , lty=2 )
lines( w.seq , pred.ci[1,] , lty=2 )
lines( w.seq , pred.ci[2,] , lty=2 )
```

(c) What aspects of the model fit concern you? Describe the kinds
of assumptions you would change, if any, to improve the model.
You don’t have to write any new code. Just explain what the model
appears to be doing a bad job of, and what you hypothesize would
be a better model.
Probability

```{r}
sample <- c("W", "L", "W", "W", "W", "L", "W", "L", "W")
W <- sum(sample=="W") # number of W observed
L <- sum(sample=="L") # number of L observed
p <- c(0, 0.25, 0.5, 0.75, 1) # proportions W
ways <- sapply(p, function(q) (q*4)^W * ((1-q)*4)^L)
prob <- ways/sum(ways)
cbind(p, ways, prob)
```

# function to toss a globe covered p by water N times
```{r}
sim_globe <- function(p=0.7, N=9) {
    sample(c("W","L"), size=N, replace=TRUE, prob=c(p,1-p))
}
```

```{r}
sim_globe()
replicate(sim_globe(p=0.5, N=9), n=10)
```

# Test the simulation on extreme settings
```{r}
sim_globe(p=1, N=11)
sum(sim_globe(p=0.5, N=le4) == "W") / le4
```