---
title: "Exercise 2"
author: "Rona Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Exercises 2

## 1.
The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. 
Provide predicted heights and 89% compatibility intervals for each of these individuals. 
That is, fill in the table below, using model-based predictions.

individual weight expected height 89% interval
1 46.95
2 43.72
3 64.78
4 32.59
5 54.63

Fit the model with quap , to provide a quadratic approximation of the posterior.

quap gives a fast, approximate Bayesian inference by fitting a Gaussian curve to the posterior near its peak,
instead of using full samplping like MCMC - Markov Chain Monte Carlo (ulam)

```{r}
library(rethinking)
data("Howell1")
d <- Howell1
d2 <- d[d$age >= 18, ] # only adults
xbar <- mean(d2$weight) # center weight
m <- quap( # fit model with quadratic approximation
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + b * (weight - xbar),
        a ~ dnorm(178, 20),
        b ~ dnorm(0, 10),
        sigma ~ dunif(0, 50)
    ),
    data = d2
)
precis(m)
```

        mean   sd   5.5%  94.5%
a     154.60 0.27 154.17 155.03
b       0.91 0.04   0.84   0.97
sigma   5.07 0.19   4.77   5.38

Then produce samples from the quadratic approximate posterior.
You could use mvrnorm directly, or the convenient extract.samples function:

```{r}
post <- extract.samples(m)
str(post)
```

'data.frame':   10000 obs. of  3 variables:
 $ a    : num  155 155 154 155 155 ...
 $ b    : num  0.85 0.89 0.879 0.861 0.871 ...
 $ sigma: num  5.3 5.03 5.05 4.92 5.2 ...
 - attr(*, "source")= chr "quap posterior: 10000 samples from m"

Now we want to plug the weights in the table into this model, and then average over the posterior to compute predictions for each individual‘s height. 
The question is ambiguous as to whether it wants m only or rather the distribution of individual height measurements (using s). 
We'll show the harder approach, that uses s and simulates individual heights. 
Also, we won't use link or sim, but it's fine if you do.

For the first individual, the code might look like this:
```{r}
# simulates 100-thousand heights, using the samples from the posterior and an assumed weight of 46.95 kg
y <- rnorm( 1e5 , post$a + post$b*( 46.95 - xbar ) , post$sigma )
# computes the average of these simulated heights
mean(y)
# computes the 89% compatibility interval of height
PI(y,prob=0.89)
```
      5%      94%
148.2392 164.5227 

To do the above for each row in the table, you can just replace the weight each time. 
We're going to write a function and use the R function sapply:
```{r}
f <- function( weight ) {
y <- rnorm( 1e5 , post$a + post$b*( weight - xbar ) , post$sigma )
return( c( mean(y) , PI(y,prob=0.89) ) )
}
weight_list <- c(46.95,43.72,64.78,32.59,54.63)
result <- sapply( weight_list , f )
```

```{r}
rtab <- cbind( weight_list , t( result ) )
colnames(rtab) <- c("weight","height","5%","94%")
rtab
```
    weight   height       5%      94%
[1,]  46.95 156.3580 148.2479 164.4669
[2,]  43.72 153.4423 145.3433 161.5515
[3,]  64.78 172.5117 164.2790 180.7596
[4,]  32.59 143.3783 135.1666 151.4932
[5,]  54.63 163.3309 155.1738 171.4928


## 2.
Select out all the rows in the Howell1 data with ages below 18 years of age. 
If you do it right, you should end up with a new data frame with 192 rows in it.

(a) Fit a linear regression to these data, using quap. 
Present and interpret the estimates. 
For every 10 units of increase in weight, how much taller does the model predict a child gets?

First make a new data frame with just the non-adults in it:
```{r}
library(rethinking)
data("Howell1")
d <- Howell1
d3 <- d[d$age < 18, ] # only <18 yrs old
str(d3)
```

'data.frame':   192 obs. of  4 variables:
 $ height: num  121.9 105.4 86.4 129.5 109.2 ...
 $ weight: num  19.6 13.9 10.5 23.6 16 ...
 $ age   : num  12 8 6.5 13 7 17 16 11 17 8 ...
 $ male  : int  1 0 0 1 0 1 0 1 0 1 ...

find the quadratic approximate posterior.
```{r}
xbar <- mean( d3$weight )
m <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - xbar ) ,
        a ~ dnorm( 178 , 20 ),
        b ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d3 )
precis(m)
```
        mean   sd   5.5%  94.5%
a     108.38 0.61 107.41 109.36
b       2.72 0.07   2.61   2.83
sigma   8.44 0.43   7.75   9.13

The estimates suggest that the MAP coefficient for weight is 2.7.
This implies that for a unit change of 1kg of weight, we predict an average of 2.7cm of increase in height.

(b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. 
Superimpose the MAP regression line and 89% interval for the mean. 
Also superimpose the 89% interval for predicted heights.
```{r}
post <- extract.samples( m )
w.seq <- seq(from=1,to=45,length.out=50)
mu <- sapply( w.seq , function(z) mean( post$a + post$b*(z-xbar) ) )
mu.ci <- sapply( w.seq , function(z)
PI( post$a + post$b*(z-xbar) , prob=0.89 ) )
pred.ci <- sapply( w.seq , function(z)
PI( rnorm( 10000 , post$a + post$b*(z-xbar) , post$sigma) , 0.89 ) )
```

```{r}
plot( height ~ weight , data=d3 ,
col=col.alpha("slateblue",0.5) , cex=0.5 )
lines( w.seq , mu )
lines( w.seq , mu.ci[1,] , lty=2 )
lines( w.seq , mu.ci[2,] , lty=2 )
lines( w.seq , pred.ci[1,] , lty=2 )
lines( w.seq , pred.ci[2,] , lty=2 )
```

(c) What aspects of the model fit concern you? Describe the kinds
of assumptions you would change, if any, to improve the model.
You don’t have to write any new code. Just explain what the model
appears to be doing a bad job of, and what you hypothesize would
be a better model.
Probability

```{r}
sample <- c("W", "L", "W", "W", "W", "L", "W", "L", "W")
W <- sum(sample=="W") # number of W observed
L <- sum(sample=="L") # number of L observed
p <- c(0, 0.25, 0.5, 0.75, 1) # proportions W
ways <- sapply(p, function(q) (q*4)^W * ((1-q)*4)^L)
prob <- ways/sum(ways)
cbind(p, ways, prob)
```

        p ways       prob
[1,] 0.00    0 0.00000000
[2,] 0.25   27 0.02129338
[3,] 0.50  512 0.40378549
[4,] 0.75  729 0.57492114
[5,] 1.00    0 0.00000000

# function to toss a globe covered p by water N times
```{r}
sim_globe <- function(p=0.7, N=9) {
    sample(c("W","L"), size=N, replace=TRUE, prob=c(p,1-p))
}
```

```{r}
sim_globe()
replicate(sim_globe(p=0.5, N=9), n=10)
```

# Test the simulation on extreme settings
```{r}
sim_globe(p=1, N=11)
sum(sim_globe(p=0.5, N=le4) == "W") / le4
```