---
title: "Assignment: AI-assisted Code Reviews"
author: "Xingrong Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Assignment: AI-assisted Code Reviews

### Setup
```{r setup, message=FALSE, warning=FALSE}
# knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# Load libraries
library(dagitty)
library(ggdag)
library(ggplot2)
library(rethinking)
library(dplyr)
library(tidyverse)
library(skimr)
# theme_set(theme_dag_blank())
```

Make sure using the same **R session** that has:
```{r}
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

### Data Inspection
```{r}
# Helps understand data structure
reviewers <- read.csv("Assignments/reviewers.csv")
reviews <- read.csv("Assignments/reviews.csv")
str(reviewers)
str(reviews)
head(reviewers)
head(reviews)
```

### Modeling Strategy
- Count data -> Poisson / Negative Binomial
- Continuous outcome -> Normal / Gamma
- Binary outcome -> Binomial (logit link)
Match the maximum entropy distribution conssitent with the data's constraints.

### Model Building Steps
1. Define causal assumptions and draw DAGs.
2. Identify adjustment sets for estimating the effect of AI usage on approval rates.
3. Prepare data: merge datasets, create key variables, standardize predictors.
4. Descriptive statistics and visualizations.
5. Statistical modeling: choose model structure based on DAGs.
6. Specify priors and conduct prior predictive checks.
7. Train models and assess convergence.
8. Compare models using LOO/WAIC.

```{r}
# Fit model using MCMC
ml <- ulam(
  alist(
    result ~ dnorm( mu, sigma ),
    mu <- a + b * x,
    a ~ dnorm( 0, 1 ),
    b ~ dnorm( 0, 1 ),
    sigma ~ dexp(1)
  ), data=mydata, chains=4, cores=4
)
```

### Compare Models
```{r}
compare(ml1, ml2, ml3)
```

### Posterior Predictive Checks
```{r}
precis(m1)
postcheck(m1)
pp_check(m1)
```

### Interpret Results


## Causal Assumptions
Outcomes:
- `extent (E)`: Number of characters in the code review.
- `quality (Q)`: Subjective quality of the code review as assessed by experts (where 1 is bad and 5 good quality).
- `result (R)`:Whether the code review contributed to making a decision on the pull request (1) or not (0).

Predictors:
- `age (G)`: Age of the reviewer in number of years.
- `skill (S)`: Skill of a reviewer measured on a continuous scale.
- `used.cr.technology (A)`: Whether the reviewer uses AI-support during code review (1) or not(0).
- `complexity (C)`: Relative complexity of the code change to be reviewed. 

### DAG 1:
Selection + Confounding + Mediation
- `age` -> `skill` -> {`extent`, `quality`} (older reviewers are usually more experienced and skilled)
- `age` -> `used.cr.technology` (older reviewers are less likely to adopt AI)
- `complexity` -> {`extent`, `quality`} (more complex code changes require longer reviews and harder to review well)
- `used.cr.technology` -> {`extent`, `quality`, `result` (aversion)} (AI-assisted can change length and clarity of reviews)
- `extent` -> `quality` -> `result`
- {`extent`, `quality`} -> `result` (more concise and higher quality reviews are more likely to have higher approval rates)
Aversion: `used.cr.technology` → `result` (direct negative path, “categorically discarded more often”)

dag {
bb="0,0,1,1"
AI [exposure,pos="0.100,0.500"]
Age [adjusted,pos="0.200,0.200"]
Complexity [adjusted,pos="0.600,0.500"]
Extent [outcome,pos="0.600,0.200"]
Quality [outcome,pos="0.600,0.800"]
Result [outcome,pos="0.900,0.500"]
Skill [pos="0.400,0.200"]
AI -> Extent
AI -> Quality
Age -> AI
Age -> Skill
Complexity -> Extent
Complexity -> Quality
Extent -> Result
Quality -> Result
Skill -> Extent
Skill -> Quality
}

dag {
bb="0,0,1,1"
AI [pos="0.100,0.500"]
Age [pos="0.200,0.200"]
Complexity [pos="0.600,0.500"]
Extent [pos="0.600,0.800"]
Quality [pos="0.600,0.200"]
Result [pos="0.900,0.500"]
Skill [pos="0.400,0.200"]
AI -> Extent
Age -> AI
Age -> Skill
Complexity -> Extent
Complexity -> Quality
Extent -> Quality
Extent -> Result
Quality -> Result
Skill -> Quality
}


Predictors (independent variables): `used.cr.technology`
Confounders: `age`, `skill`, `complexity`
Mediators: `extent`, `quality`
Outcomes (dependent variables): `result`

Therefore, 
- `age`, `skill`, and `complexity` are confounders that influence both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.
- `extent` and `quality` are mediators that lie on the causal path between the predictor `used.cr.technology` and the outcome `result`.

This DAG supports total vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.

age ─▶ skill ─▶ extent ─▶ result
 │             └▶ quality ─▶ result
 └────────▶ used.cr.technology ─▶ extent, quality ─▶ result
complexity ─▶ extent, quality, result

```{r dag1, fig.height=5, fig.width=7}
dag1 <- dagitty("dag {
    age -> skill -> extent
    age -> skill -> quality
    age -> used.cr.technology
    complexity -> extent
    complexity -> quality
    used.cr.technology -> extent
    used.cr.technology -> quality
    extent -> result
    quality -> result
}")

ggdag(dag1, text = FALSE, layout = "nicely") +
geom_dag_edges(edge_colour = "grey30") +
geom_dag_point(shape = 21, size = 16, fill = "#8ECAE6", colour = "#4E9AC3", stroke = 0.8) +
geom_dag_text(aes(label = name), size = 4, colour = "black") +
labs(title = "DAG 1: Selection + Confounding + Mediation") +
coord_equal()
```

### DAG 2：
Skill-driven; AI mainly a proxy for skill & complexity
adoption + Confounding + Mediation
- `age` -> `skill` (older reviewers are usually more experienced and skilled)
- `skill` -> `used.cr.technology` (more skilled reviewers are usually curious or confident enough to try new tools.)
- `skill` -> `extent` & `quality` (more skilled reviewers write more concise and higher quality reviews)
- `complexity` -> `extent` & `quality` (more complex code changes require longer reviews and harder to review well)
- `extent` & `quality` -> `result`,  (more concise and higher quality reviews are more likely to have higher approval rates)
`used.cr.technology` -> `result` (weak/zero) direct path to Result; main AI influence is mediated by Extent/Quality

Allow that `used.cr.technology` is a proxy for skill and complexity, and the main influence of AI on `result` is mediated by `extent` and `quality`.
most of the observed difference between AI and non-AI reviews is due to differences in skill and complexity, not a direct effect of AI.
Skill is a confounder that influences both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.

The apparent correlation between “AI users have lower approval rate” might not mean AI causes worse results.
It could mean “AI users are younger, less experienced, or handle harder reviews.”

After adjusting for those confounders (Skill, Complexity, Age), the direct effect of AI may disappear.

This DAG supports total vs adjusted vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
- Adjusted effect of `used.cr.technology` on `result` conditioning on confounders `age`, `skill`, and `complexity` but not mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.

age ─▶ skill ─▶ used.cr.technology ─▶ extent ─▶ result
 │                     │                └▶ quality ─▶ result
 │                     └────────────────────────────────▶ result (weak)
complexity ─────────────────────────▶ extent, quality, result

```{r dag2, fig.height=5, fig.width=7}
dag2 <- dagitty("dag {
    G -> S
    S -> A
    S -> E
    S -> Q
    C -> E
    C -> Q
    E -> R
    Q -> R
    A -> R
}")

ggdag(dag2, text = FALSE, layout = "nicely") +
geom_dag_edges(edge_colour = "grey30") +
geom_dag_point(shape = 21, size = 16, fill = "#BDEFB6", colour = "#69BE63", stroke = 0.8) +
geom_dag_text(aes(label = name), size = 4, colour = "black") +
labs(title = "DAG 2: Skill-driven; AI as proxy for skill & complexity") +
coord_equal()
```

```{r adjustment-sets}
adjustmentSets(dag1, exposure = "A", outcome = "R")
adjustmentSets(dag2, exposure = "A", outcome = "R")
```

Viability discussion. 
Compare which back-door paths need blocking and what each adjustment set implies (e.g., in DAG 1, include Age, Skill, Complexity to estimate AI’s effect; if you include Extent/Quality you move from total to direct effect). 
Tie your reasoning to the collider/mediator logic practiced in E3 (information, K–L, thinking carefully about model vs. data) and E5 (mediation pipes).

---

## Descriptive Statistics & Visualizations

Follow the standardize/visualize patterns from E2/E4:
Summaries and correlations (extent/quality/complexity/age/skill/AI/result).
Visuals: scatter (Extent vs Complexity), outcome by AI (box/scatter), approval rates by AI.
Normalize extent/quality where helpful; keep raw scales visible for interpretability.

```{r data-prep}
# reviewers <- read.csv("Assignments/reviewers.csv")
# str(reviewers) # check format
# reviewers <- as.list(reviewers) # make to a list so we can add alpha below

# reviews <- read.csv("Assignments/reviews.csv")
# str(reviews) # check format
# reviews <- as.list(reviews) # make to a list so we can add alpha below

# Merge datasets
# df <- merge(reviews, reviewers, by = "reviewer_id")
# glimpse(df)

# Rename variables for Stan compatibility
# df <- df %>%
#     rename(
#         used_cr_technology = used.cr.technology,
#     )

# Standardize predictors
# df <- df %>%
#     mutate(
#         age_s = scale(age),
#         skill_s = scale(skill),
#         complexity_s = scale(complexity),
#         extent_s = scale(extent),
#         quality_s = scale(quality)
#     )
```

Load files:
```{r descriptives, fig.height=5, fig.width=7}
reviewers <- read_csv("Assignments/reviewers.csv")
reviews   <- read_csv("Assignments/reviews.csv")

# inspect
head(reviewers)
head(reviews)
```
  reviewer.id   age skill used.cr.technology
        <dbl> <dbl> <dbl>              <dbl>
1           1    51  4.25                  0
2           2    55  3.74                  0
3           3    49  4.46                  0
4           4    38  3.83                  0
5           5    44  4.18                  0
6           6    27  2.66                  1

  reviewer.id review.id complexity result extent quality
        <dbl>     <dbl>      <dbl>  <dbl>  <dbl>   <dbl>
1           1       623      0.367      1    229       3
2           1      1381      0.490      1    184       3
3           1       673     -1.40       0    126       3
4           1       136      0.545      0    183       3
5           1        35     -0.882      0    169       3
6           1       749      0.919      1    179       4

Merge tha tables:
```{r}
df <- reviews %>%
  left_join(reviewers, by = "reviewer.id")

# quick structure check
glimpse(df)
```
Rows: 2,000
Columns: 9
$ reviewer.id        <dbl> 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, …
$ review.id          <dbl> 623, 1381, 673, 136, 35, 749, 626, 620, 495, 1136, …
$ complexity         <dbl> 3.669993e-01, 4.899597e-01, -1.396837e+00, 5.453157…
$ result             <dbl> 1, 1, 0, 0, 0, 1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 1, 0, …
$ extent             <dbl> 229, 184, 126, 183, 169, 179, 299, 257, 167, 320, 1…
$ quality            <dbl> 3, 3, 3, 3, 3, 4, 3, 4, 4, 4, 3, 3, 3, 4, 3, 3, 5, …
$ age                <dbl> 51, 51, 51, 51, 51, 51, 51, 55, 55, 55, 55, 55, 55,…
$ skill              <dbl> 4.254235, 4.254235, 4.254235, 4.254235, 4.254235, 4…
$ used.cr.technology <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …

Create Key Variables:
```{r}
df <- df %>%
    mutate(
        ai_user = as.integer(used.cr.technology),
        approved = as.integer(result == 1)
    )
```

Model:
```{r}
m1_total <- glm(result ~ usd.cr.technology + age + skill + complexity,
                data = df, family = binomial(link = "logit"))
```

Direct effect:
```{r}
m2_direct <- glm(result ~ usd.cr.technology + age + skill + complexity + extent + quality,
                 data = df, family = binomial(link = "logit"))
```

Descriptive Statistics:
Overall summaries:
```{r}
skim(df %>% select(extent, quality, complexity, age, skill))
```
── Variable type: numeric ──────────────────────────────────────────────────────
  skim_variable n_missing complete_rate     mean     sd     p0     p25      p50
1 extent                0             1 191.     72.2   34     139     180
2 quality               0             1   3.41    0.642  1       3       3
3 complexity            0             1  -0.0383  0.990 -3.06   -0.705  -0.0308
4 age                   0             1  40.1     9.83  20      33      41
5 skill                 0             1   3.61    0.802  0.419   3.22    3.74  
      p75   p100 hist
1 236     538    ▃▇▃▁▁
2   4       5    ▁▁▇▇▁
3   0.628   3.36 ▁▅▇▃▁
4  48      62    ▃▅▇▅▂
5   4.17    4.92 ▁▁▂▇▆

Groupwise summaries by AI usage:
```{r}
df %>%
  group_by(ai_user) %>%
  summarise(
    mean_extent = mean(extent, na.rm = TRUE),
    mean_quality = mean(quality, na.rm = TRUE),
    mean_complexity = mean(complexity, na.rm = TRUE),
    approval_rate = mean(approved, na.rm = TRUE),
    n = n()
  )
```
  ai_user mean_extent mean_quality mean_complexity approval_rate     n
    <int>       <dbl>        <dbl>           <dbl>         <dbl> <int>
1       0        180.         3.48         -0.0605         0.589   917
2       1        200.         3.35         -0.0195         0.404  1083

Visualizations:
Extent vs Complexity:
```{r}
ggplot(df, aes(x = complexity, y = extent)) +
  geom_point(alpha = 0.5) +
  labs(title = "Extent vs Complexity",
       x = "Complexity of code review",
       y = "Extent (characters)") +
  theme_minimal()
```

Quality by AI usage
```{r}
ggplot(df, aes(x = factor(ai_user, labels = c("No AI", "AI")), y = quality)) +
  geom_boxplot() +
  labs(title = "Quality by AI usage",
       x = "Used code review technology",
       y = "Quality rating") +
  theme_minimal()
```

Approval Rate by AI usage
```{r}
df %>%
  group_by(ai_user) %>%
  summarise(approval_rate = mean(approved, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(ai_user, labels = c("No AI", "AI")), y = approval_rate)) +
  geom_col(fill = "steelblue") +
  ylim(0,1) +
  labs(title = "Approval rate by AI usage",
       x = "Used code review technology",
       y = "Approval rate") +
  theme_minimal()
```

Correlation matrix
```{r}
df %>%
  select(extent, quality, complexity, age, skill, ai_user, approved) %>%
  cor(use = "complete.obs") %>%
  round(3)
```
           extent quality complexity    age  skill ai_user approved
extent      1.000  -0.261      0.692 -0.272 -0.337   0.140    0.379
quality    -0.261   1.000     -0.144  0.371  0.496  -0.098    0.148
complexity  0.692  -0.144      1.000 -0.026 -0.019   0.021    0.441
age        -0.272   0.371     -0.026  1.000  0.776  -0.413    0.104
skill      -0.337   0.496     -0.019  0.776  1.000  -0.345    0.109
ai_user     0.140  -0.098      0.021 -0.413 -0.345   1.000   -0.184
approved    0.379   0.148      0.441  0.104  0.109  -0.184    1.000

```{r}
vars <- df %>% 
  select(extent, quality, complexity, age, skill, used.cr.technology, approved)

cor_mat <- cor(vars, use = "complete.obs")
round(cor_mat, 2)
```

```{r}
# Compute significance (p-values)
cor_test <- cor.mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], conf.level = conf.level)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
      lowCI.mat[i, j] <- lowCI.mat[j, i] <- tmp$conf.int[1]
      uppCI.mat[i, j] <- uppCI.mat[j, i] <- tmp$conf.int[2]
    }
  }
  list(p = p.mat, lowCI = lowCI.mat, uppCI = uppCI.mat)
}

p.mat <- cor_test(vars)$p
```

```{r}
library(ggcorrplot)

ggcorrplot(cor_mat,
           hc.order = TRUE,          # hierarchical clustering order
           type = "lower",           # show only lower triangle
           lab = TRUE,               # show correlation values
           p.mat = p.mat,            # add significance info
           sig.level = 0.05,         # mark significant correlations
           insig = "blank",          # hide non-significant
           colors = c("red", "white", "blue"),
           title = "Correlation Matrix of Key Variables",
           lab_size = 3) +
  theme_minimal()
```

```{r}
cor_filtered <- cor_mat
cor_filtered[abs(cor_filtered) < 0.3] <- NA

ggcorrplot(cor_filtered,
           type = "lower",
           lab = TRUE,
           colors = c("red", "white", "blue"),
           title = "Strong Correlations (|r| > 0.3)") +
  theme_minimal()

```
Table X presents the correlation matrix of the key variables.
The results reveal strong positive associations between complexity and extent (r = 0.69) and between age and skill (r = 0.78).
Skill also correlates with quality (r = 0.50), suggesting that more experienced reviewers tend to produce higher-quality feedback.
Negative correlations between ai_user and both age (r = –0.41) and skill (r = –0.35) indicate that younger, less-experienced reviewers are more likely to use AI technology.
The approval rate is moderately correlated with extent (r = 0.38) and complexity (r = 0.44), while showing a weak negative correlation with AI usage (r = –0.18).
These findings support the inclusion of complexity, extent, age, skill, and AI usage as key variables in the causal models of Task 3.



The dataset contains 200 reviewers and 2000 reviews.
The mean extent is X, with higher values for complex reviews.
Non-AI reviewers have an average approval rate of ≈ 0.59, while AI users’ approval rate is ≈ 0.40.
Skill and age correlate positively with quality (r ≈ 0.5 and 0.37 respectively), indicating more experienced reviewers write higher-quality comments.
Complexity and extent are strongly correlated (r ≈ 0.69).


## Statistical Modeling
### Option (a): 1 DAG → 1 model per outcome
Extent (continuous): linear model (or Student-t if heavy tails):
Extent_𝑖 ∼ 𝑁 (𝜇_𝑖, 𝜎), 𝜇_𝑖 = 𝛼 + 𝛽_AI * AI_𝑖 + 𝛽_𝑐 * Complexity_𝑖 + 𝛽_𝑎 * Age_𝑖 + 𝛽_𝑠 * Skill_𝑖.
(E2 shows linear regression workflow; E4 suggests adding varying intercepts by reviewer if you want a multilevel version.)
 
Quality (ordered 1–5): ordered logistic/probit:

Pr(Quality_𝑖 ≤ 𝑘) = logit*(−1) * (𝜏_𝑘 − 𝜂_𝑖), 𝜂_𝑖 = 𝛾_0 + 𝛾_AI * AI_𝑖 + 𝛾_𝑐 * Complexity_𝑖 + 𝛾_𝑎 * Age_𝑖 + 𝛾_𝑠 * Skill_𝑖.

(Ordered categorical modeling is exactly what E6 trains, including priors on cutpoints.)


Result (binary): logistic:

Result_𝑖 ∼ Bernoulli(𝑝_𝑖), logit(𝑝_𝑖) = 𝛿_0 + 𝛿_AI * AI_𝑖 + 𝛿_𝑐 * Complexity_𝑖 + 𝛿_𝑎 * Age_𝑖 + 𝛿_𝑠 * Skill_𝑖 + [𝛿_𝑒 * Extent_𝑖 + 𝛿_𝑞 * Quality_𝑖],

bracketed terms included when you target the direct (not total) effect of AI (E5).
Add random intercepts for reviewer.id to acknowledge repeated reviews (E4).

### Option (b): 1 outcome across ≥2 DAGs (recommended: Result)

Model R₁ (DAG 1, total effect): logit(Result) ~ AI + Age + Skill + Complexity.

Model R₂ (DAG 1, direct effect): logit(Result) ~ AI + Age + Skill + Complexity + Extent + Quality.
Compare R₁ vs R₂ (WAIC/LOO). If 𝛿_AI shrinks a lot in R₂, the effect is mediated by Extent/Quality (or by the aversion path you posit). (E5 gives the template for this comparison.)

## Priors & Prior Predictive Checks
Use standardized inputs to simplify priors (E2/E4/E6 style):

Intercepts (logit/ordered): Normal(0,1) or Normal(−1,1) when you believe baseline approval < 50% (see NWO example).

Slopes: Normal(0, 0.5) or Normal(0, 1) on standardized predictors (regularizing as in Wines2012).

Cutpoints (ordered quality): ordered priors (e.g., Normal with increasing constraints). (E6)

Group SDs (if multilevel): half-Normal(0, 1) on random-effect scales (E4).
Run prior predictive checks to ensure plausible ranges before training (E4/E6).

## Assess Training & Compare Models
Convergence (R-hat, trace plots), effective sample sizes (if Bayesian; E4).

LOO/WAIC to compare candidate models and decide which DAG mapping predicts best (E4/E7).

## Conclusions
Report causal statements with the adjustment set tied to your DAG.

Contrast total vs direct effect of AI on Result.

Discuss bias/aversion (direct AI → Result path) and whether differences remain after adjusting for Extent/Quality/Complexity.
This is exactly what the assignment asks you to synthesize at the end.