---
title: "Assignment: AI-assisted Code Reviews"
author: "Xingrong Zong"
date: "`r Sys.Date()`"
output: html_document
---

# Assignment: AI-assisted Code Reviews

### Modeling Strategy
- Count data -> Poisson / Negative Binomial
- Continuous outcome -> Normal / Gamma
- Binary outcome -> Binomial (logit link)
Match the maximum entropy distribution conssitent with the data's constraints.

### Model Building Steps
1. Define **causal assumptions** and draw **DAGs**.
2. Identify **adjustment sets** for estimating the effect of AI usage on approval rates.
Ajustment set is a set of variables to condition on in order to block all backdoor paths from the treatment (AI usage) to the outcome (approval rate), 
thus isolating the causal effect of the treatment on the outcome.
3. Prepare data: merge datasets, create key variables, standardize predictors.
4. **Descriptive** statistics and **visualizations**.
5. Statistical modeling: choose  structure based on DAGs.
```{r}
# Fit model using MCMC
ml <- ulam(
  alist(
    result ~ dnorm( mu, sigma ),
    mu <- a + b * x,
    a ~ dnorm( 0, 1 ),
    b ~ dnorm( 0, 1 ),
    sigma ~ dexp(1)
  ), data=mydata, chains=4, cores=4
)
```
1. Identify outcome variables.
2. Justigy predictors suing causal reasoning (DAG) and maximum entropy / GLM principles.
3. Specify priors and conduct prior predictive checks.
4. Train models and assess convergence with rethinking (ulam) or Stan / brms.
5.  Compare models using LOO/WAIC.
```{r}
compare(ml1, ml2, ml3)
```
1.  Discuss posterior predictions, regulatization, and causla interpretation.
```{r}
precis(m1)
postcheck(m1)
pp_check(m1)
```

---

## Causal Assumptions
Variables:
- `extent`: Number of characters in the code review.
- `quality`: Subjective quality of the code review as assessed by experts (where 1 is bad and 5 good quality).
- `result`:Whether the code review contributed to making a decision on the pull request (1) or not (0).
- `age`: Age of the reviewer in number of years.
- `skill`: Skill of a reviewer measured on a continuous scale.
- `used.cr.technology`: Whether the reviewer uses AI-support during code review (1) or not(0).
- `complexity`: Relative complexity of the code change to be reviewed. 

### DAG 1:
Selection + Confounding + Mediation
- `age` -> `skill` -> {`extent`, `quality`} (older reviewers are usually more experienced and skilled)
- `age` -> `used.cr.technology` (older reviewers are less likely to adopt AI)
- `complexity` -> {`extent`, `quality`} (more complex code changes require longer reviews and harder to review well)
- `used.cr.technology` -> {`extent`, `quality`, `result` (aversion)} (AI-assisted can change length and clarity of reviews; direct negative path, “categorically discarded more often”)
- {`extent`, `quality`} -> `result` (more concise and higher quality reviews are more likely to have higher approval rates)

dag {
  bb="0,0,1,1"
  AI [exposure,pos="0.100,0.500"]
  Age [adjusted,pos="0.200,0.200"]
  Complexity [adjusted,pos="0.600,0.500"]
  Extent [outcome,pos="0.600,0.200"]
  Quality [outcome,pos="0.600,0.800"]
  Result [outcome,pos="0.900,0.500"]
  Skill [pos="0.400,0.200"]
  AI -> Extent
  AI -> Quality
  AI -> Result
  Age -> AI
  Age -> Skill
  Complexity -> Extent
  Complexity -> Quality
  Extent -> Result
  Quality -> Result
  Skill -> Extent
  Skill -> Quality
}

Predictors (independent variables): `used.cr.technology`
Confounders: `age`, `skill`, `complexity`
Mediators: `extent`, `quality`
Outcomes (dependent variables): `result`

Therefore, 
- {`age`, `skill`, `complexity`} are confounders that influence both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.
- {`extent`, `quality`} are mediators that lie on the causal path between the predictor `used.cr.technology` and the outcome `result`.

This DAG supports total vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
  - Exposure (What want to study): `used.cr.technology`.
  - Minimal **Adjustment set** (variables that bias if ignore) for confounder `age`.
  - Adjust for pre-treatment **precision variables** (helpful extra variables that make estimates sharper) {`skill`, `complexity`}.
    - Improve efficiency, reduce uncertainty, get narrower posterior intervals, stabilize the model.
- Adjusted effect of `used.cr.technology` on `result` conditioning on confounders `age`, `skill`, and `complexity` but not mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.


### DAG 2:
Skill-driven; AI mainly a proxy for skill & complexity
adoption + Confounding + Mediation
- `age` -> `skill` (older reviewers are usually more experienced and skilled)
- `skill` -> `used.cr.technology` (more skilled reviewers are usually curious or confident enough to try new tools.)
- `skill` -> {`extent`, `quality`} (more skilled reviewers write more concise and higher quality reviews)
- `complexity` -> {`extent`, `quality`} (more complex code changes require longer reviews and harder to review well)
- {`extent`, `quality`} -> `result`,  (more concise and higher quality reviews are more likely to have higher approval rates)
- `used.cr.technology` -> `result` (weak/zero) direct path to Result; main AI influence is mediated by Extent/Quality

dag {
  bb="0,0,1,1"
  AI [pos="0.100,0.500"]
  Age [pos="0.200,0.200"]
  Complexity [pos="0.600,0.500"]
  Extent [pos="0.600,0.800"]
  Quality [pos="0.600,0.200"]
  Result [pos="0.900,0.500"]
  Skill [pos="0.400,0.200"]
  AI -> Extent
  Age -> AI
  Age -> Skill
  Complexity -> Extent
  Complexity -> Quality
  Extent -> Quality
  Extent -> Result
  Quality -> Result
  Skill -> Quality
}

Allow that `used.cr.technology` is a proxy for skill and complexity, and the main influence of AI on `result` is mediated by `extent` and `quality`.
most of the observed difference between AI and non-AI reviews is due to differences in skill and complexity, not a direct effect of AI.
Skill is a confounder that influences both the predictor `used.cr.technology` and the outcomes `extent`, `quality`, and `result`.

The apparent correlation between “AI users have lower approval rate” might not mean AI causes worse results.
It could mean “AI users are younger, less experienced, or handle harder reviews.”

After adjusting for those confounders (Skill, Complexity, Age), the direct effect of AI may disappear.

This DAG supports total vs adjusted vs direct effect reasoning:
- Total effect of `used.cr.technology` on `result` without conditioning on mediators `extent` and `quality`.
- Adjusted effect of `used.cr.technology` on `result` conditioning on confounders `age`, `skill`, and `complexity` but not mediators `extent` and `quality`.
- Direct effect of `used.cr.technology` on `result` conditioning on mediators `extent` and `quality` and confounders `age`, `skill`, and `complexity`.
---

## Descriptive Statistics & Visualizations
### Load Data
```{r descriptives, message=FALSE, warning=FALSE}
library(rethinking)
library(tidyverse)

reviewers <- read.csv("Assignments/reviewers.csv")
reviews <- read.csv("Assignments/reviews.csv")
str(reviewers)
str(reviews)
head(reviewers)
head(reviews)
```
'data.frame':   200 obs. of  4 variables:
 $ reviewer.id       : int  1 2 3 4 5 6 7 8 9 10 ...
 $ age               : int  51 55 49 38 44 27 45 45 35 42 ...
 $ skill             : num  4.25 3.74 4.46 3.83 4.18 ...
 $ used.cr.technology: int  0 0 0 0 0 1 0 1 1 1 ...

 'data.frame':   2000 obs. of  6 variables:
 $ reviewer.id: int  1 1 1 1 1 1 1 2 2 2 ...
 $ review.id  : int  623 1381 673 136 35 749 626 620 495 1136 ...
 $ complexity : num  0.367 0.49 -1.397 0.545 -0.882 ...
 $ result     : int  1 1 0 0 0 1 1 1 0 1 ...
 $ extent     : int  229 184 126 183 169 179 299 257 167 320 ...
 $ quality    : int  3 3 3 3 3 4 3 4 4 4 ...

 reviewer.id age    skill used.cr.technology
1           1  51 4.254235                  0
2           2  55 3.739723                  0
3           3  49 4.457738                  0
4           4  38 3.827274                  0
5           5  44 4.184495                  0
6           6  27 2.657038                  1

reviewer.id review.id complexity result extent quality
1           1       623  0.3669993      1    229       3
2           1      1381  0.4899597      1    184       3
3           1       673 -1.3968369      0    126       3
4           1       136  0.5453157      0    183       3
5           1        35 -0.8819071      0    169       3
6           1       749  0.9185690      1    179       4

```{r dag1}
library(dagitty)
dag1 <- dagitty("dag{
  age -> skill
  skill -> extent
  skill -> quality
  age -> used.cr.technology
  complexity -> extent
  complexity -> quality
  used.cr.technology -> extent
  used.cr.technology -> quality
  used.cr.technology -> result
  extent -> result
  quality -> result
}")
```

| Outcome   | Type of variable               | Suggested distribution     | Link function    | Predictors                                         |
| --------- | ------------------------------ | -------------------------- | ---------------- | -------------------------------------------------- |
| `extent`  | continuous (length/time)       | Normal(μ, σ)               | identity         | `skill`, `complexity`, `used.cr.technology`, `age` |
| `quality` | continuous or bounded score    | Normal(μ, σ) or Beta(α, β) | identity / logit | `skill`, `complexity`, `used.cr.technology`, `age` |
| `result`  | binary (approved vs discarded) | Binomial(n, p)             | logit            | `extent`, `quality`, `used.cr.technology`          |

### Merge Data
```{r data-prep}
library(dplyr)
library(skimr)

join_key <- intersect(names(reviewers), names(reviews))
if (!length(join_key)) stop("No shared key between reviewers and reviews. Please align id column(s).")
df <- reviews %>% left_join(reviewers, by = join_key)
# df <- reviews %>% left_join(reviewers, by = "reviewer_id")

glimpse(df)
```
Rows: 2,000
Columns: 9
$ reviewer.id        <int> 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, …
$ review.id          <int> 623, 1381, 673, 136, 35, 749, 626, 620, 495, 1136, …
$ complexity         <dbl> 3.669993e-01, 4.899597e-01, -1.396837e+00, 5.453157…
$ result             <int> 1, 1, 0, 0, 0, 1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 1, 0, …
$ extent             <int> 229, 184, 126, 183, 169, 179, 299, 257, 167, 320, 1…
$ quality            <int> 3, 3, 3, 3, 3, 4, 3, 4, 4, 4, 3, 3, 3, 4, 3, 3, 5, …
$ age                <int> 51, 51, 51, 51, 51, 51, 51, 55, 55, 55, 55, 55, 55,…
$ skill              <dbl> 4.254235, 4.254235, 4.254235, 4.254235, 4.254235, 4…
$ used.cr.technology <int> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …

### Standardize & Prepare Data List
```{r}
# Rename to match DAG (adapt to your real column names!)
df <- df %>%
  rename(
    age        = age,                 # reviewer age
    skill      = skill,               # reviewer skill/expertise
    complexity = complexity,          # change complexity
    used       = used.cr.technology,  # AI usage (0/1)
    extent     = extent,              # review length / duration
    quality    = quality,             # quality score
    result     = result               # approval (0/1)
  )

glimpse(df)
```

- Precents mnuerical instability -> smoother MCMC sampling
- Makes priors meaningful -> Normal(0, 0.5) now corresponds to a reasonable effect size
- Easier interpretation -> slope = effect of 1 SD increase in the predictor
- Helps avoid divergent rtrnsitions in HMC/Stan
```{r}
# Coerce binaries and standardize predictors (z = (x-mean)/sd)
df <- df %>%
  mutate(
    used   = as.integer(used),
    result = as.integer(result),
    z_age        = as.numeric(scale(age)),
    z_skill      = as.numeric(scale(skill)),
    z_complexity = as.numeric(scale(complexity)),
    z_extent     = as.numeric(scale(extent)),
    z_quality    = as.numeric(scale(quality))
  )

glimpse(df)
```
Rows: 2,000
Columns: 14
$ reviewer.id  <int> 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, …
$ review.id    <int> 623, 1381, 673, 136, 35, 749, 626, 620, 495, 1136, 1258, …
$ complexity   <dbl> 3.669993e-01, 4.899597e-01, -1.396837e+00, 5.453157e-01, …
$ result       <int> 1, 1, 0, 0, 0, 1, 1, 1, 0, 1, 1, 0, 0, 0, 0, 1, 0, 1, 1, …
$ extent       <int> 229, 184, 126, 183, 169, 179, 299, 257, 167, 320, 162, 66…
$ quality      <int> 3, 3, 3, 3, 3, 4, 3, 4, 4, 4, 3, 3, 3, 4, 3, 3, 5, 4, 4, …
$ age          <int> 51, 51, 51, 51, 51, 51, 51, 55, 55, 55, 55, 55, 55, 49, 4…
$ skill        <dbl> 4.254235, 4.254235, 4.254235, 4.254235, 4.254235, 4.25423…
$ used         <int> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …
$ z_age        <dbl> 1.1120172, 1.1120172, 1.1120172, 1.1120172, 1.1120172, 1.…
$ z_skill      <dbl> 0.8027060, 0.8027060, 0.8027060, 0.8027060, 0.8027060, 0.…
$ z_complexity <dbl> 0.40957361, 0.53382569, -1.37279220, 0.58976325, -0.85245…
$ z_extent     <dbl> 0.53001807, -0.09341851, -0.89695898, -0.10727265, -0.301…
$ z_quality    <dbl> -0.6372565, -0.6372565, -0.6372565, -0.6372565, -0.637256…

### Descriptive Stats & Visuals
- Helps detect skewness (important for choosing Poisson vs Negative-Binomial)
- Shows whether variables need transformation (log, standardization)
- Shows if outcome is balanced or imbalanced, which affects model choice
- Helps confirm assumptions in DAG
```{r}
library(ggplot2)

summary(df[,c("age","skill","complexity","extent","quality")])
table(used = df$used, result = df$result)

skim(df)
```
   age            skill          complexity           extent     
 Min.   :20.00   Min.   :0.4191   Min.   :-3.06063   Min.   : 34.0  
 1st Qu.:33.00   1st Qu.:3.2154   1st Qu.:-0.70517   1st Qu.:139.0  
 Median :41.00   Median :3.7388   Median :-0.03076   Median :180.0  
 Mean   :40.07   Mean   :3.6108   Mean   :-0.03832   Mean   :190.7
 3rd Qu.:48.00   3rd Qu.:4.1678   3rd Qu.: 0.62845   3rd Qu.:236.0  
 Max.   :62.00   Max.   :4.9244   Max.   : 3.35950   Max.   :538.0
    quality
 Min.   :1.000  
 1st Qu.:3.000
 Median :3.000
 Mean   :3.409
 3rd Qu.:4.000  
 Max.   :5.000

 result
used   0   1
   0 377 540
   1 645 438

Bar Plot of AI usage
```{r}
ggplot(df, aes(x=factor(used))) + geom_bar() + labs(x="AI used", y="Count")
```

Bar Plot of Approval Result
```{r}
ggplot(df, aes(x=factor(result))) + geom_bar() + labs(x="Approval", y="Count")
```

Scatter Plots to check relationships complexity vs extent
```{r}
ggplot(df, aes(z_complexity, z_extent)) + geom_point(alpha=.3) + geom_smooth(method="lm", se=FALSE)
```

Scatter Plots to check relationships skill vs quality
```{r}
ggplot(df, aes(z_skill, z_quality)) + geom_point(alpha=.3) + geom_smooth(method="lm", se=FALSE)
```

Alternatives
```{r}
ggplot(df_plot, aes(x = factor(used_cr_tech), y = quality)) + geom_boxplot()
ggplot(df_plot, aes(x = complexity, y = extent, color = factor(result))) + geom_point(alpha = 0.4)
ggplot(df_plot, aes(x = skill, y = extent, color = factor(used_cr_tech))) + geom_point(alpha = 0.4)
ggplot(df_plot, aes(x = age, y = quality, color = factor(used_cr_tech))) + geom_point(alpha = 0.4)
```

## Priors & Prior Predictive Checks 
```{r}
library(rethinking)

data_list <- list(
  extent = df$z_extent,
  quality = df$quality,
  result = df$result,
  skill = standardize(df$skill),
  complexity = standardize(df$complexity),
  used_cr_tech = df$used.cr.technology,
  age = standardize(df$age)
)
glimpse(data_list)
```

### Extent model (Gaussian on log-extent)
```{r}
# EXTENT
dl_extent <- list(
  extent = df$z_extent,
  age = df$z_age, 
  skill = df$z_skill,
  complexity = df$z_complexity, 
  used = df$used
)
```

```{r}
m_extent_prior <- ulam(
  alist(
    extent ~ dnorm(mu, sigma),
    mu <- a + b_age*age + b_skill*skill + b_complexity*complexity + b_used*used,
    a ~ dnorm(0, 1),
    c(b_age, b_skill, b_complexity, b_used) ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = dl_extent, 
  chains=1, 
  iter=500, 
  sample=TRUE
)
```

```{r}
# Prior predictive: simulate with extract.prior() then link()
# draws from the *prior* (ignoring likelihood)
prior_ext <- extract.prior(m_extent_prior)

# Pick some "typical" covariate rows for visualization, e.g. mean 0 for all z's:
newdat_ext <- list(
  age        = rep(0, 100),
  skill      = rep(0, 100),
  complexity = rep(0, 100),
  used       = rep(0, 50) %>% c(rep(1, 50))   # 50 without AI, 50 with AI
)

mu_prior_ext <- link(m_extent_prior, post = prior_ext, data = newdat_ext)
# Turn linear predictor into simulated outcomes
y_prior_ext  <- sim(m_extent_prior, post = prior_ext, data = newdat_ext)
```
```{r}
# 4) Inspect the distribution of prior predictive outcomes
summary(as.vector(y_prior_ext))
hist(as.vector(y_prior_ext), breaks = 40, main = "Prior predictive for standardized extent")
```

dat_ext <- list(
  N = nrow(dat),
  used = dat$used,
  complexity = dat$complexity_z,
  skill = dat$skill_z,
  age = dat$age_z
)

m_ext_prior <- ulam(
  alist(
    extent_log ~ dnorm(mu, sigma),
    mu <- a + bU*used + bC*complexity + bS*skill + bA*age,
    a  ~ dnorm(0, 1.5),
    c(bU, bC, bS, bA) ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = c(dat_ext, list(extent_log = dat$extent_log)),
  chains = 1, iter = 1000, sample = "prior"
)
# Check implied extent distribution:
postcheck(m_ext_prior)  # should broadly cover log-length range

# 2) Quality model (ordered logistic)
# rethinking uses ordered_logistic() with cutpoints
dat_qual <- list(
  N = nrow(dat),
  K = 5,
  used = dat$used,
  complexity = dat$complexity_z,
  skill = dat$skill_z,
  age = dat$age_z,
  extent = dat$extent_z,
  quality = dat$quality_int
)

m_qual_prior <- ulam(
  alist(
    quality ~ ordered_logistic(eta, cutpoints),
    eta <- a + bU*used + bC*complexity + bS*skill + bA*age + bE*extent,
    a ~ dnorm(0, 1.5),
    c(bU, bC, bS, bA, bE) ~ dnorm(0, 0.5),
    cutpoints ~ dnorm(0,1)
  ),
  data = dat_qual,
  chains = 1, iter = 1000, sample = "prior"
)
# No postcheck for ordered directly, but check range of eta:
precis(m_qual_prior)

# 3) Result model (binomial/logit)
dat_res <- list(
  N = nrow(dat),
  result = dat$result,
  used = dat$used,
  extent = dat$extent_z,
  quality = dat$quality_z
)

m_res_prior <- ulam(
  alist(
    result ~ dbinom(1, p),
    logit(p) <- a + bQ*quality + bE*extent + bU*used,
    a ~ dnorm(0, 1.5),
    c(bQ, bE, bU) ~ dnorm(0, 0.5)
  ),
  data = dat_res,
  chains = 1, iter = 1000, sample = "prior"
)
precis(m_res_prior)

# ====== FIT MODELS (posterior) ======
```{r}
m_extent <- ulam(
  alist(
    extent ~ dnorm(mu, sigma),
    mu <- a + b_skill*skill + b_complexity*complexity +
           b_ai*used_cr_tech + b_age*age,
    a ~ dnorm(0, 1),
    c(b_skill, b_complexity, b_ai, b_age) ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data=data_list, chains=4, cores=4
)

# m_ext <- ulam(
#   alist(
#     extent_log ~ dnorm(mu, sigma),
#     mu <- a + bU*used + bC*complexity + bS*skill + bA*age,
#     a  ~ dnorm(0, 1.5),
#     c(bU, bC, bS, bA) ~ dnorm(0, 0.5),
#     sigma ~ dexp(1)
#   ),
#   data = c(dat_ext, list(extent_log = dat$extent_log)),
#   chains = 4, cores = 4, iter = 4000, log_lik = TRUE
# )
```

```{r}
m_quality <- ulam(
  alist(
    quality ~ dnorm(mu, sigma),
    mu <- a + b_skill*skill + b_complexity*complexity +
           b_ai*used_cr_tech + b_age*age,
    a ~ dnorm(0, 1),
    c(b_skill, b_complexity, b_ai, b_age) ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data=data_list, chains=4, cores=4
)

# m_qual <- ulam(
#   alist(
#     quality ~ ordered_logistic(eta, cutpoints),
#     eta <- a + bU*used + bC*complexity + bS*skill + bA*age + bE*extent,
#     a ~ dnorm(0, 1.5),
#     c(bU, bC, bS, bA, bE) ~ dnorm(0, 0.5),
#     cutpoints ~ dnorm(0,1)
#   ),
#   data = dat_qual,
#   chains = 4, cores = 4, iter = 4000, log_lik = TRUE
# )
```

```{r}
m_result <- ulam(
  alist(
    result ~ dbinom(1, p),
    logit(p) <- a + b_extent*extent + b_quality*quality + b_ai*used_cr_tech,
    a ~ dnorm(0, 1),
    c(b_extent, b_quality, b_ai) ~ dnorm(0, 1)
  ), data=data_list, chains=4, cores=4
)

# m_res <- ulam(
#   alist(
#     result ~ dbinom(1, p),
#     logit(p) <- a + bQ*quality + bE*extent + bU*used,
#     a ~ dnorm(0, 1.5),
#     c(bQ, bE, bU) ~ dnorm(0, 0.5)
#   ),
#   data = dat_res,
#   chains = 4, cores = 4, iter = 4000, log_lik = TRUE
# )
```

# ====== DIAGNOSTICS ======
```{r}
precis(m_extent)
precis(m_quality)
precis(m_result)

precis(m_extent, 2);  trankplot(m_extent);  traceplot(m_extent)
precis(m_quality, 2); trankplot(m_quality); traceplot(m_quality)
precis(m_result, 2);  trankplot(m_result);  traceplot(m_result)

compare(m_extent, m_quality, m_result, func=PSIS)
```

# quick helpers:
max(abs(coef(m_ext)$Rhat - 1))
max(abs(coef(m_qual)$Rhat - 1))
max(abs(coef(m_res)$Rhat - 1))

# ====== POSTERIOR PREDICTIVE CHECKS ======
postcheck(m_ext)   # should resemble log-extent distribution
# For ordered quality, inspect cutpoints & fitted category probs
post <- extract.samples(m_qual)
summary(post$cutpoints)

# ====== (Optional) MODEL VARIANTS + PSIS-LOO ======
# e.g., compare extent models with/without skill to see confounding
m_ext_noskill <- ulam(
  alist(
    extent_log ~ dnorm(mu, sigma),
    mu <- a + bU*used + bC*complexity + bA*age,
    a ~ dnorm(0,1.5),
    c(bU,bC,bA) ~ dnorm(0,0.5),
    sigma ~ dexp(1)
  ),
  data = c(dat_ext, list(extent_log = dat$extent_log)),
  chains = 4, cores = 4, iter = 3000, log_lik = TRUE
)

compare(m_ext, m_ext_noskill, func = PSIS)  # lower = better predictive