---
title: "Exercise 5"
author: "Rona Zong"
date: "`r Sys.Date()`"
output: html_document
---

# 1.
G - Gender
D - Discipline
A - Awards

G -> A
G -> D -> A

G: predictor (independent variable)
D: mediator
A: outcome (dependent variable)

Total effect of G on A without conditioning on mediator D.

```{r setup, include=FALSE}
library(rethinking)
data(NWOGrants)
d <- NWOGrants
dat_list <- list(
  awards = as.integer(d$awards),
  apps = as.integer(d$applications),
  gid = ifelse( d$gender == "m", 1L, 2L))
ml_total <- ulam(
    alist(
        awards ~ binomial(apps, p),
        logit(p) <- a[gid], 
        a[gid] ~ normal(-1, 1) # Only gender is included as predictor -> this captures total effect (direct + indirect)
    ), data = dat_list, chains = 4)
precis(ml_total, 2)
```
      mean   sd  5.5% 94.5% rhat ess_bulk
a[1] -1.54 0.06 -1.64 -1.43    1  1235.63
a[2] -1.74 0.08 -1.87 -1.61    1  1366.46

```{r}
post <- extract.samples(ml_total)
diff <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis(list(diff=diff))
```

'data.frame': 2000 obs. of 1 variables:
     mean   sd 5.5% 94.5%  histogram
diff 0.03 0.01 0.01  0.05 ▁▁▂▅▇▇▃▁▁▁

Direct effect of G on A conditioning on mediator D.

```{r}
dat_list$disc <- as.integer(d$discipline)
m1_direct <- ulam(
    alist(
        awards ~ binomial( apps , p ),
        logit(p) <- a[gid] + d[disc],
        a[gid] ~ normal(-1,1),
        d[disc] ~ normal(0,1) # control for discipline D
    ),data=dat_list , chains=4 , cores=4 , cmdstan=TRUE )
precis(m1_direct ,2)
```
      mean   sd  5.5% 94.5% rhat ess_bulk
a[1] -1.38 0.33 -1.90 -0.86 1.03   163.10
a[2] -1.51 0.33 -2.04 -0.99 1.03   167.78
d[1]  0.36 0.37 -0.23  0.95 1.02   216.74
d[2]  0.03 0.35 -0.53  0.58 1.02   186.68
d[3] -0.20 0.35 -0.76  0.34 1.02   182.57
d[4] -0.24 0.38 -0.85  0.36 1.01   221.25
d[5] -0.30 0.34 -0.84  0.25 1.02   179.12
d[6]  0.02 0.37 -0.57  0.59 1.02   202.16
d[7]  0.33 0.39 -0.29  0.95 1.02   215.57
d[8] -0.42 0.34 -0.96  0.11 1.03   173.50
d[9] -0.17 0.36 -0.76  0.39 1.02   195.51


```{r}
post <- extract.samples(m1_direct)
diff_a <- post$a[,1] - post$a[,2]
precis( list( diff_a=diff_a ) )
```
'data.frame': 2000 obs. of 1 variables:
       mean   sd  5.5% 94.5% histogram
diff_a 0.14 0.11 -0.03  0.31  ▁▂▅▇▅▁▁▁